A gap-filling approach to evaluation of machine translation for assimilation

### 2015-02-10 2015-eamt-assim/2015-eamt-assim.tex
import stuffs
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58816 72bbbca6-d526-0410-a7d9-f06f51895060

As suggested by [Church, Hovy 1993], modern machine translation (MT) systems may
be divided into two broad categories according to their purpose: post-editing systems, e.g.
Google Translate 1 , and assimilation systems, e.g. Apertium 2 . Both kinds may be
systematically evaluated, either to control for quality in the development process or to
compare the systems. The evaluation methods, however, differ depending on the type of
the system. The methods are described in detail in [Church, Hovy 1993].
In this work I perform the evaluation of an assimilation MT system called Apertium.
Apertium is an open-source community-maintained project that provides translation for
39 language pairs. The goal of this study is to assess whether machine translation
performed by this system assists users in understanding the gist of texts in an unknown
language. The evaluation is performed in three separate experiments for the following
translation directions: Basque to Spanish, English to Kazakh, and Tatar to Russian.
I build on the methodology established in [Trosterund, Unhammer 2012] and
\cite{oregan2013}, which are the works describing previous evaluations of
Apertium language pairs. [O’Regan, Forcada 2013] proposes the gap-filling task as a
means of assessing participant’s understanding of text. The same methodology is used in
this work and described in detail in the section below.

>>>>

As suggested by \newcite{church93}, modern machine translation (MT) systems may
be divided into two broad categories according to their purpose: post-editing systems, e.g.
Google Translate 1 , and assimilation systems, e.g. Apertium 2 . Both kinds may be
systematically evaluated, either to control for quality in the development process or to
compare the systems. The evaluation methods, however, differ depending on the type of
the system. The methods are described in detail in \newcite{church93}.
In this work I perform the evaluation of an assimilation MT system called Apertium.
Apertium is an open-source community-maintained project that provides translation for
39 language pairs. The goal of this study is to assess whether machine translation
performed by this system assists users in understanding the gist of texts in an unknown
language. The evaluation is performed in three separate experiments for the following
translation directions: Basque to Spanish, English to Kazakh, and Tatar to Russian.
I build on the methodology established in \newcite{trosterud12} and
\newcite{oregan2013}, which are the works describing previous evaluations of
Apertium language pairs. \newcite{oregan2013} proposes the gap-filling task as a
means of assessing participant’s understanding of text. The same methodology is used in
this work and described in detail in the section below.
